Something about correlation functions

### 2021-06-25 paper-correlations.tex
Add section about testing

\documentclass[reprint,amsmath,amssymb,aps,pre]{revtex4-1}
% Latex
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Xelatex
%\usepackage{polyglossia}
%\usepackage{fontspec}
%\setdefaultlanguage{english}
%\setmainfont{DejaVu Sans}
%\setsansfont{DejaVu Sans}
%\setmonofont{DejaVu Sans Mono}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algpseudocode}

\definecolor{light-gray}{gray}{0.95}
\newcommand{\code}[1]{\colorbox{light-gray}{\texttt{#1}}}
% When cleveref fails to do its job
\newcommand{\apref}[1]{Appendix \ref{#1}}
\DeclareMathOperator*{\argmin}{arg\,min}

@@

These correlation functions are calculated using two different approaches (when
applicable). The first approach is implemented in the module \code{Directional}
and the second approach is in the module \code{Map}. The former approach (called
the scan approach) calculates correlation functions along specified directions
which can include axial and diagonal directions. The results of computations can
be provided to the user separately for each requested direction or averaged over
multiple directions. The latter approach (called the map approach) calculates a
``correlation map'' which holds the values of a correlation function in all
directions. The scan and map approaches are described in \cref{scansec} and
\cref{mapsec} respectively.

@@

Two-point function is so heaviliy used internally in the package so we propose a
bit more general definition for it:
\begin{equation*}
  S_2^{\chi}(\bm{r}) = \langle \chi(\bm{x}, \bm{x} + \bm{r}) \rangle
\end{equation*}
where $\chi(\bm{x})$ is some arbitrary indicator (or membership) function. and
$\bm{x} \in \mathbb{R}^{2n}$ where $n$ is dimensionality of the input array. The
vector $\bm{x}$ is constructed as $(\bm{x_1}, \bm{x_2})$ where $\bm{x_1}$ and
$\bm{x_2}$ are the coordinates of the both ends of a line segment thrown into
the input array.

@@

The algorithm for calculation of the array \code{success} for
$F_{sv}^{(i)}(\bm{r})$ for one-dimensional slice is very similar to the previous
one:
\begin{algorithmic}[1]
  \Procedure{surfvoid}{$array, phase, periodic$}
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edge \gets \text{Detect edges in $ind$}$
    \State $indices \gets \text{Array of all indices in $array$}$
    \State $\chi_1(x) \gets edge[x]$
    \State $\chi_2(x) \gets array[x] = 0$ \Comment{$0$ is a value which designates the
      void phase}
    \State \textbf{return} $s2sep(indices, \chi_1, \chi_2, periodic)$
  \EndProcedure
\end{algorithmic}

@@

As you can see this function cannot be obtained from one-dimensional slices. We
calculate it using distance transform. For each point $\bm{x}$ distance
transform $\mathcal{D}(\bm{x})$ is defined as
\begin{equation*}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \argmin\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right.
\end{equation*}

@@

\subsection{Chord length function}
The chord length function $p^{(i)}(\bm{r})d\bm{r}$ is defined as the probability of
finding a chord of length between $|\bm{r}|$ and $|\bm{r} + d\bm{r}|$ and
paralled to $\bm{r}$ in phase $i$. A chords is a line segment between
intersections of an infinitely long line with the two-phase interface. This
function, again, can be calculated by cutting slices from a multidimensional
array. The algorithm is as follows:
\begin{algorithmic}[1]
  \Procedure{chordlength}{$array, phase, nbins$}
    \State $lengths \gets [\quad]$
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edgephase \gets maximum(array) + 1$ \Comment{``bogus'' phase for the
      interface of the phase $phase$}
    \State $edge \gets edgephase \cdot [\text{Detect edges in $ind$}]$ \Comment{We
      use distance transform here}
    \State $combo = min(edge + array, edgephase)$ \Comment{Combine the original
      array and the interface for the phase $phase$}
    \ForAll{$direction \in \{\text{Requested directions}\}$}
      \ForAll{$slice \in \{\text{Slices parallel to $direction$}\}$}
        \State $len \gets 0$
        \State $startonedge \gets false$
        \ForAll{$x \in slice$}
          \If{$x = edgephase$}
            \Comment{Edge found}
            \If{$len > 1 \land startonedge$}
              \State $lengths \gets (len-1):lengths$
            \EndIf
            \State $len \gets 0$
            \State $startonedge \gets true$
          \ElsIf{$x \ne phase$}
            \State $startonedge \gets false$
          \EndIf
          \State $len \gets len + 1$
        \EndFor
      \EndFor
    \EndFor
    \State Bin elements in $lengths$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}

@@

\section{Verifying our algorithms}
To verify our package we use random datasets of overlapping spheres with radii
$R$ and centers placed in points generated by Poisson point process with a
parameter $\lambda$. For these datasets correlation functions can be given in
closed-form expressions for one-, two- and three-dimensional cases. For the sake
of testing we obtain a collection of spheres in a square (the number of
dimensions $n = 2$) or cube ($n = 3$) with the side $s$ using the following
algorithm:
\begin{algorithmic}[1]
  \Procedure{spheres}{$s, n, R, \lambda$}
    \State $nspheres \gets poisson(\lambda s^n)$ \Comment{$poisson(\lambda)$
      generates random numbers with Poisson distribution with parameter
      $\lambda$.}
    \State Generate $nspheres$ random points which are uniformly distributes in
    a cube ($n = 3$) or square ($n = 2$) with a side $s$
    \State Draw spheres of radius $R$ in those points. Store the result in an
    array $output$.
    \State \textbf{return} $output$
  \EndProcedure
\end{algorithmic}

>>>>

\documentclass[reprint,amsmath,amssymb,aps,pre]{revtex4-1}
% Latex
\usepackage[english]{babel}
\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
% Xelatex
%\usepackage{polyglossia}
%\usepackage{fontspec}
%\setdefaultlanguage{english}
%\setmainfont{DejaVu Sans}
%\setsansfont{DejaVu Sans}
%\setmonofont{DejaVu Sans Mono}
\usepackage{bm}
\usepackage{cleveref}
\usepackage{xcolor}
\usepackage{algpseudocode}
\usepackage{graphicx}
\usepackage{subcaption}

\definecolor{light-gray}{gray}{0.95}
\newcommand{\code}[1]{\colorbox{light-gray}{\texttt{#1}}}
% When cleveref fails to do its job
\newcommand{\apref}[1]{Appendix \ref{#1}}

@@

These correlation functions are calculated using two different approaches (when
applicable). The first approach is implemented in the module \code{Directional}
and the second approach is in the module \code{Map}. The former approach (called
the scan approach) calculates correlation functions along specified directions
which can include axial and diagonal directions. The results of computations can
be provided to the user separately for each requested direction or averaged
across multiple directions. The latter approach (called the map approach)
calculates a ``correlation map'' which holds the values of a correlation
function in all directions. The scan and map approaches are described in
\cref{scansec} and \cref{mapsec} respectively.

@@

Two-point function is so heaviliy used internally in the package so we propose a
bit more general definition for it:
\begin{equation*}
  S_2^{\chi}(\bm{r}) = \langle \chi(\bm{x}, \bm{x} + \bm{r}) \rangle
\end{equation*}
where $\chi(\bm{x})$ is some arbitrary indicator (or membership) function and
$\bm{x} \in \mathbb{R}^{2n}$ where $n$ is dimensionality of the input array. The
vector $\bm{x}$ is constructed as $(\bm{x_1}, \bm{x_2})$ where $\bm{x_1}$ and
$\bm{x_2}$ are the coordinates of the both ends of a line segment thrown into
the input array.

@@

The algorithm for calculation of the array \code{success} for
$F_{sv}^{(i)}(\bm{r})$ for one-dimensional slice is very similar to the previous
one:
\begin{algorithmic}[1]
  \Procedure{surfvoid}{$array, phase, periodic$}
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edge \gets \text{Detect edges in $ind$}$
    \State $indices \gets \text{Array of all indices in $array$}$
    \State $\chi_1(x) \gets edge[x]$
    \State $\chi_2(x) \gets array[x] = 0$ \Comment{$0$ is a value reserved for the
      void phase}
    \State \textbf{return} $s2sep(indices, \chi_1, \chi_2, periodic)$
  \EndProcedure
\end{algorithmic}

@@

As you can see this function cannot be obtained from one-dimensional slices. We
calculate it using distance transform. For each point $\bm{x}$ distance
transform $\mathcal{D}(\bm{x})$ is defined as
\begin{equation*}
  \mathcal{D}(\bm{x})= \left\{
  \begin{array}{ll}
    0 & \quad \bm{x} \in \text{solid phase} \\
    \min\limits_{y \in \text{solid phase}} \rho(\bm{x},\bm{y}) & \quad \text{otherwise}
  \end{array}
\right.
\end{equation*}

@@

\subsection{Chord length function}
The chord length function $p^{(i)}(\bm{r})d\bm{r}$ is defined as the probability of
finding a chord of length between $|\bm{r}|$ and $|\bm{r} + d\bm{r}|$ and
paralled to $\bm{r}$ in phase $i$. A chords is a line segment between
intersections of an infinitely long line with the interface of the phase
$i$. This function, again, can be calculated by cutting slices from a
multidimensional array. The algorithm is as follows:
\begin{algorithmic}[1]
  \Procedure{chordlength}{$array, phase, nbins$}
    \State $lengths \gets [\quad]$
    \State $ind \gets array = phase$ \Comment{$array$ is compared with $phase$
      element-wise}
    \State $edgephase \gets maximum(array) + 1$ \Comment{``bogus'' phase for the
      interface of the phase $phase$}
    \State $edge \gets edgephase \cdot [\text{Detect edges in $ind$}]$ \Comment{We
      use distance transform here}
    \State $combo = min(edge + array, edgephase)$ \Comment{Combine the original
      array and the interface for the phase $phase$}
    \ForAll{$direction \in \{\text{Requested directions}\}$}
      \ForAll{$slice \in \{\text{Slices parallel to $direction$}\}$}
        \State $len \gets 0$
        \State $startonedge \gets false$
        \ForAll{$x \in slice$}
          \If{$x = edgephase$}
            \Comment{Edge found}
            \If{$len > 1 \land startonedge$}
              \State $lengths \gets (len-1):lengths$
            \EndIf
            \State $len \gets 0$
            \State $startonedge \gets true$
          \ElsIf{$x \ne phase$}
            \State $startonedge \gets false$
          \EndIf
          \State $len \gets len + 1$
        \EndFor
      \EndFor
    \EndFor
    \State Bin elements in $lengths$ to $nbins$ bins to compute a histogram $H$
    \State \textbf{return} $H$
  \EndProcedure
\end{algorithmic}

@@

\section{Verifying our algorithms}
To verify our package we use random datasets of overlapping disks (the number of
dimensions $n = 2$) or balls ($n = 3$) with radii $R$ and centers placed in
points generated by Poisson point process with a parameter $\lambda$. For these
datasets correlation functions can be given in closed-form expressions for one-,
two- and three-dimensional cases. For the sake of testing we obtain a collection
of disks (resp. balls) in a square (resp. cube) with the side $s$ using the
following algorithm:
\begin{algorithmic}[1]
  \label{testdata}
  \Procedure{testdata}{$s, n, R, \lambda$}
    \State $nballs \gets poisson(\lambda s^n)$ \Comment{$poisson(\lambda)$
      generates random numbers with Poisson distribution with parameter
      $\lambda$.}
    \State Generate $nballs$ random points which are uniformly distributes in
    a cube ($n = 3$) or square ($n = 2$) with a side $s$
    \State Draw balls (disks) of radius $R$ in those points. Store the result in
    an array $output$.
    \State \textbf{return} $output$
  \EndProcedure
\end{algorithmic}

In two-dimensional case we test functions $L_2(r)$, $S_2(r)$, $F_{ss}(r)$ and
$F_{sv}(r)$ using parameters $R = 100$, $\lambda = 4\cdot10^{-5}$ and
$s = 5000$. The correlation functions are calculated in two axial directions
with periodic boundary conditions and the result is averaged across the
directions. Instead of taking one realization of Poisson process we take 15
realizations and average results over them. The result can be seen on
\cref{fig:2d}.

For pore size function $P(r)$ and chord length function $p(\bm{r})$ a slightly
different approach is required. We take a single realization of Poisson point
process with the same parameters $R$ and $\lambda$ in a square with the side
$s = 20000$ (resp. $s = 12000$) and calculate histograms for Pore size
(resp. Chord length) function. Chord length function is calculated in two axial
directions. The theoretical values are integrated over bin ranges and compared
with the number of elements in those bins. The results are presented on
\cref{fig:pscl}.

In three-dimensional case the approach for testing is still the same but with
different parameters to speed up calculations. We choose $R = 20$,
$\lambda = 3\cdot10^{-5}$ and $s = 500$. The correlation functions are
calculated in three axial directions with periodic boundary conditions and the
result is also averaged across the directions as in two-dimensional
case. We take six realizations of Poisson process for averaging the results
which are presented on \cref{fig:3d}.

Pore size and chord length functions are calculated using the side of a cube
$s = 700$. The results are presented on \cref{fig:pscl}.

The results for three-dimensional case are seemingly worse than for
two-dimensional case. This can be explained by a reduced number of realizations
of Poisson process and smaller value of $s$.

\onecolumngrid
\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/l2-2d.png}
    \caption[]{{\small Lineal-path correlation function}}
    \label{fig:l2-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/s2-2d.png}
    \caption[]{{\small Two-point correlation function}}
    \label{fig:s2-2d}
  \end{subfigure}
  \vskip\baselineskip
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ss-2d.png}
    \caption[]{{\small Surface-surface correlation function}}
    \label{fig:ss-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/sv-2d.png}
    \caption[]{{\small Surface-void correlation function}}
    \label{fig:sv-2d}
  \end{subfigure}
  \caption[]{\small A comparison of calculated values of $L_2(r)$, $S_2(r)$,
    $F_{ss}(r)$ and $F_{sv}(r)$ with theoretical values for overlapping disks.}
  \label{fig:2d}
\end{figure*}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/l2-3d.png}
    \caption[]{{\small Lineal-path correlation function}}
    \label{fig:l2-3d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/s2-3d.png}
    \caption[]{{\small Two-point correlation function}}
    \label{fig:s3-3d}
  \end{subfigure}
  \vskip\baselineskip
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ss-3d.png}
    \caption[]{{\small Surface-surface correlation function}}
    \label{fig:ss-3d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/sv-3d.png}
    \caption[]{{\small Surface-void correlation function}}
    \label{fig:sv-3d}
  \end{subfigure}
  \caption[]{\small A comparison of calculated values of $L_2(r)$, $S_2(r)$,
    $F_{ss}(r)$ and $F_{sv}(r)$ with theoretical values for overlapping balls.}
  \label{fig:3d}
\end{figure*}

\begin{figure*}[t]
  \centering
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ps-2d.png}
    \caption[]{{\small Pore size correlation function (2D)}}
    \label{fig:ps-2d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/cl-2d.png}
    \caption[]{{\small Chord length correlation function (2D)}}
    \label{fig:cl-2d}
  \end{subfigure}
  \vskip\baselineskip
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/ps-3d.png}
    \caption[]{{\small Pore size correlation function (3D)}}
    \label{fig:ps-3d}
  \end{subfigure}
  \hfill
  \begin{subfigure}[b]{0.475\textwidth}
    \centering
    \includegraphics[width=\textwidth]{images/cl-3d.png}
    \caption[]{{\small Chord length correlation function (3D)}}
    \label{fig:cl-3d}
  \end{subfigure}
  \caption[]{\small A comparison of calculated values of $P(r)$ and $p(r)$
    with theoretical values for overlapping disks (\cref{fig:ps-2d} and
    \cref{fig:cl-2d}) and balls(\cref{fig:ps-3d} and \cref{fig:cl-3d}).}
  \label{fig:pscl}
\end{figure*}
\twocolumngrid
